---
title: "pull data together"
author: "David Hughes"
date: "2/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## pulling the FGFP microbiome, metabolome, and clinical trait data together

### Related Study: FGFP mt -> metabolite association analysis for a breast cancer study

### For: Kaitlin Wade, Tim Robinson, and Grace Edmunds

# Read in paramater file

```{r}
library(tidyverse)
source("parameters/pfile.sh")
```

# Read in the data

```{r}
#############################
## 1) linker file
#############################
linker = read.table(linker_file, header = TRUE, as.is = TRUE, sep = "\t")

#############################
## 2) primary clinical file
#############################
clinical_data = read.table(clinical_trait_file, header = TRUE, as.is = TRUE, sep = "\t")


#############################
## 6) metabolite data
#############################
metabolite_data = read.table(metabolite_data_file, header = TRUE, as.is = TRUE, sep = "\t")
met_feature_data = read.table(metabolite_feature_file, header = TRUE, as.is = TRUE, sep = "\t")
met_sample_data = read.table(metabolite_sample_file, header = TRUE, as.is = TRUE, sep = "\t")
################
## add VDP ids
################
m = match( rownames(metabolite_data), linker$VERC_id )
## how many metabolite samples do I not have a VDP ID for??
sum(is.na(m)) ## 58
## remove those with NO match
w = which(is.na(m))
metabolite_data = metabolite_data[-w,]
met_sample_data = met_sample_data[-w,]
## perform match again
m = match( rownames(metabolite_data), linker$VERC_id )
rownames(metabolite_data) = linker$fgfp_id[m]

#############################
## 7) microbiome data
#############################
mt_data = read.table(microbial_trait_file, header = TRUE, as.is = TRUE, sep = "\t")
mt_batch_data = read.table(microbial_lab_batch_file, header = TRUE, as.is = TRUE, sep = "\t")
mt_phylogenetic_data = read.table(microbial_phylogenetics_file, header = TRUE, as.is = TRUE, sep = "\t")
## edit identifier
mt_batch_data$SAMPLE_NAME = gsub("VDP.","VDP.0", mt_batch_data$SAMPLE_NAME)

```

# Vector of all unique VDP IDs

```{r}
## make a vector of all the sample VDP IDs
vdp_ids = sort( unique( c( 
                   rownames(mt_data), 
                   mt_batch_data$SAMPLE_NAME, ## microbial traits
                   rownames(metabolite_data), ## metabolite
                   rownames(clinical_data), ##  clinical  data
                   linker$fgfp_id ## linker file
                   ) ) )

vdp_ids = as.character(vdp_ids)
length( vdp_ids ) ## 3133
```

# generate uniform sample ID data frames

## metabolite data

```{r}
#####################
## Metabolite data
#####################
m = match(vdp_ids, rownames(metabolite_data) )
metabolite_data = metabolite_data[m, ]
rownames(metabolite_data) = vdp_ids

```

## microbiome data

```{r}
m = match(vdp_ids, rownames(mt_data) )
mt_data = mt_data[m, ]
rownames(mt_data) = vdp_ids

## batch data
rownames(mt_batch_data) = mt_batch_data$SAMPLE_NAME
mt_batch_data = mt_batch_data[, -1]
m = match(vdp_ids, rownames(mt_batch_data ) )
mt_batch_data = mt_batch_data[m, ]
rownames(mt_batch_data) = vdp_ids

## redefine MT data
mt_data = cbind( mt_batch_data , mt_data)
```


## clincal data

```{r}
#####################
## Clinical data
#####################
m = match(vdp_ids, rownames(clinical_data) )
clinical_data = clinical_data[m, ]
rownames(clinical_data) = vdp_ids

```


# Write data to file

```{r}
### Clinical Data
f = paste0(data_dir, "clinical_data.txt")
write.table(clinical_data, file = f, row.names = TRUE, col.names = TRUE, sep = "\t", quote = FALSE)

### Microbiome Data
f = paste0(data_dir, "microbiome_data.txt")
write.table(mt_data, file = f, row.names = TRUE, col.names = TRUE, sep = "\t", quote = FALSE)

f = paste0(data_dir, "microbiome_phylogeny.txt")
write.table(mt_phylogenetic_data, file = f, row.names = FALSE, col.names = TRUE, sep = "\t", quote = FALSE)

### Metabolite Data
f = paste0(data_dir, "metabolite_data.txt")
write.table(metabolite_data, file = f, row.names = TRUE, col.names = TRUE, sep = "\t", quote = FALSE)

f = paste0(data_dir, "metabolite_feature_data.txt")
write.table(met_feature_data, file = f, row.names = TRUE, col.names = TRUE, sep = "\t", quote = FALSE)

```

